---
title: "College tuition, diversity, and pay"
author: "Zahra Khoshmanesh, Vahid Aziziz, Saba Moeinizadeh"
date: "4/25/2020"
output: slidy_presentation
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(tidyverse)
library(lubridate)

address<-c('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2020/2020-03-10/')

tuition_cost <- readr::read_csv(paste(address,'tuition_cost.csv',sep=""))

tuition_income <- readr::read_csv(paste(address,'tuition_income.csv',sep=""))

salary_potential <- readr::read_csv(paste(address,'salary_potential.csv',sep=""))

historical_tuition <- readr::read_csv(paste(address,'historical_tuition.csv',sep=""))

diversity_school <- readr::read_csv(paste(address,'diversity_school.csv',sep=""))
```

## College tuition, diversity, and pay

- Many people are interested to know about the tuition, costs, diversity and potential salary when searching for college. 
- In this project, we want to analyze tuition costs across different states, and explore diversity in different schools. 
- The data set includes different variables such as school name, state, type of school, in-state/out-of-state tuition, group/racial/gender category, early/mid-career pay, stem percent and historical tuition information. 
- We are also interested in trends of tuition over time. We will use different summary statistics and visualizations in R to address these problems.


## Diversity dataset

```{r, echo = TRUE}

diversity_school %>% glimpse()

```

## Finding the most diverse campus per each diversity categories

```{r,echo = TRUE}
max_per_category <- diversity_school %>% 
  mutate(eroll_rate_category=(enrollment/total_enrollment)*100) %>% 
  group_by(category) %>% 
  summarise(university_name=name[which.max(eroll_rate_category)],
            state=state[which.max(eroll_rate_category)],
            enrollment_rate=eroll_rate_category[which.max(eroll_rate_category)],
            total_enroll=total_enrollment[which.max(eroll_rate_category)],
            category_enroll=enrollment[which.max(eroll_rate_category)])

max_per_category

```

## Finding the most diverse campus per each diversity categories

```{r}
max_per_category %>% 
  ggplot(aes(x=category,y=enrollment_rate,fill=university_name,label=state)) +
  geom_col() +
  geom_text(nudge_y = -3) +
  coord_flip()
```

