---
title: "College tuition, diversity, and pay"
author: "Zahra Khoshmanesh, Vahid Aziziz, Saba Moeinizadeh"
date: "4/25/2020"
output: slidy_presentation
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

## College tuition, diversity, and pay

- Many people are interested to know about the tuition, costs, diversity and potential salary when searching for college. 
- In this project, we want to analyze tuition costs across different states, and explore diversity in different schools. 
- The data set includes different variables such as school name, state, type of school, in-state/out-of-state tuition, group/racial/gender category, early/mid-career pay, stem percent and historical tuition information. 
- We are also interested in trends of tuition over time. We will use different summary statistics and visualizations in R to address these problems.

## Slide with Bullets

- Bullet 1
- Bullet 2
- Bullet 3

## Slide with R Output

```{r cars, echo = TRUE}
summary(cars)
```

## Slide with Plot

```{r pressure}
plot(pressure)
```

